---
title: "alpha_missense_plotting"
output: html_notebook
---
# code by: Matthew Howard @ UCSF
# -
# the goal of this code is to take alpha missense residue-level pathogenicity data for the enture GPCRome and create a databese of heatmaps. 
# -
# set working directory, reset to source folder on your system, this is where to put the AM data and where your heatmaps and stuctures will go
```{r}
setwd("/Users/..........") 
```
# - 
# source packages
```{r}
install.packages("tidyverse")
install.packages("bio3d")
library(tidyverse)
library(bio3d)
```
# - 
# define order of variants on heatmaps
```{r}
order_missense <- c("A", "G", "M", "V", "L", "I", "T", "S", "C", "Q", "N", "Y", "W", "F", "E", "D", "H", "K", "R", "P")
```
# -
# make function for parsing AM data, plotting heatmaps. just run ths block to init the functionss
```{r}

#process the alpha missense file
process_AM_file <- function(file) {
  # Read the TSV file
  data <- read.delim(file, header = TRUE, sep = "\t")

  # Rename the columns
  colnames(data) <- c("uniprot_id", "protein_variant", "am_pathogenicity", "am_class")

  # HGVS parsing
  data <- data %>%
    mutate(protein_variant = paste('p.(', protein_variant, ')', sep = ''))
  data$pos <- as.integer(str_extract(data$protein_variant, "\\d+"))
  data$variants <- str_sub(data$protein_variant, -2, -2)
  data$is.wt <- str_sub(data$protein_variant, 4, 4) == str_sub(data$protein_variant, -2, -2)
  
  return(data)
}
# example usage
# tshr_AF_missense <- process_AM_file("tshr_alpha_missense.tsv")

#plot a heatmap
create_heatmap <- function(gene, order, chunk_size, scores_df, score_column, variant_names = NULL, output_file = NULL) {
  
  # Check if variant_names were provided; if not, use the original variant names
  if (is.null(variant_names)) {
    variant_names <- order
  }
  
  scores_df <- scores_df %>%
    filter(variants %in% order)
  
  gene_chunks <- str_split(gene, '')[[1]]
  num_chunks <- ceiling(length(gene_chunks) / chunk_size)

  min_score <- min(scores_df[, score_column], na.rm = TRUE)
  max_score <- max(scores_df[, score_column], na.rm = TRUE)

  heatmap_plots <- list()

  for (i in 1:num_chunks) {
    start_idx <- (i - 1) * chunk_size + 1
    end_idx <- min(i * chunk_size, length(gene_chunks))

    gene_chunk <- gene_chunks[start_idx:end_idx]
    scores_chunk <- scores_df[scores_df$pos %in% c(start_idx:(start_idx + length(gene_chunk) - 1)), ]
    
    # Replace the variant names with the desired names
    scores_chunk$variants <- factor(scores_chunk$variants, levels = order, labels = variant_names)

    
    heatmap_plot <- ggplot(data = scores_chunk, aes(x = pos, y = factor(variants, level = order), fill = !!sym(score_column))) +
      geom_tile(aes(color = "grey"), linewidth = 0.2, position = "identity") +
      scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, l1 = 0.2, l3 = 0.2, p1 = 0.9, p3 = .4, p4 = 0.7, rev = FALSE, na.value = 'grey', limits = c(min_score, max_score)) +
theme(
        panel.background = element_rect(fill = "grey", size = 0.1, linetype = "solid"),
        panel.grid.major = element_line(size = 0.1, linetype = 'solid', colour = "grey"),
        panel.grid.minor = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.ticks.x.top = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(family = "mono", size = 7, angle = 0, hjust = 0.5, vjust = 1, margin = margin(t = 1)),
        axis.text = element_text(family = "mono", size = 7, color = "black"),
        axis.text.x.top = element_text(family = "mono", size = 7, angle = 0, hjust = 0.5, margin = margin(r = 0)),
        axis.text.y = element_text(family = "mono", size = 7,  margin = margin(r = 0)) ) +
      scale_x_continuous(breaks = seq(0, 1000, by = 5), expand = c(0, 0),
                         sec.axis = sec_axis(trans = ~., name = "Sequence", breaks = seq(start_idx, end_idx), labels = gene_chunk, guide = derive())) +
      coord_fixed(ratio = 1) +
      scale_color_manual(values = c(NA, 'green')) +
      labs(y = "Mutation", x = "Position") +
      geom_tile(data = subset(scores_chunk, !is.wt), aes(color = "green"), size = 0.2, position = "identity") +
      geom_tile(data = subset(scores_chunk, is.wt), aes(color = "grey"), size = 0.2, position = "identity")
    
    heatmap_plots[[i]] <- heatmap_plot
  }
  heatmap_combined <- ggarrange(plotlist = heatmap_plots, nrow = num_chunks, ncol = 1)
  ggsave(output_file, heatmap_combined, height = 10, width = 15, dpi = 600)
}
# Example usage
# create_heatmap(gene, order, chunk_size, scores_df, score_column, variant_names = c("Name1", "Name2", "Name3"), output_file = "heatmap.png")


map_scores_pdb <- function(input_pdb, mapping_scores, field, selection = NULL) {
  
  if (is.null(selection)) {
    
    selection = atom.select(input_pdb, "protein")
  }
  
  output_pdb = trim.pdb(input_pdb, selection)
  
  for (i in seq_len(dim(output_pdb$atom[1]))) {
    
    if (output_pdb$atom[i,]$resno > 0) {
      
      n = as.character(output_pdb$atom[i,]$resno)
      j = which(mapping_scores['pos'] == n)
      
      if (length(j) == 0) {
        score = 0
        
      } else {
        score = mapping_scores[j, field][[1]]
      }
      
      if (!is.na(score)) {
        
        output_pdb$atom[i,]$b = score
        
      } else {
        
        output_pdb$atom[i,]$b = 0
        
      }
    } else {
      output_pdb$atom[i,]$b = 0
    }
  }
  
  return(output_pdb)
}
```
# -
#loop over the entire GPCRome
```{r}

library(readr)

# Read the CSV file
csv_data <- read_csv("GPCRome/uniprot_seq.csv")
colnames(csv_data) <- c("uniprot_id", "gene_name", "seq")

# Set the subfolder name
subfolder_name <- "GPCRome"  # Replace with the actual subfolder name

# Construct the full path to the subfolder
subfolder_path <- file.path(getwd(), subfolder_name)

# Change the working directory to the subfolder
setwd(subfolder_path)

# Read the CSV file
csv_data <- read_csv("uniprot_seq.csv")

# Loop through each row of the CSV file and call the functions
for (i in 1:nrow(csv_data)) {
  gene_name <- csv_data$gene_name[i]  # Assuming "GeneName" is the column name with gene names
  amino_acid_seq <- csv_data$seq[i]  # Assuming "AminoAcidSeq" is the column name with sequences

  # Call process_AM_file function
  tsv_file_name <- paste0(gene_name, ".tsv")
  tsv_data <- process_AM_file(tsv_file_name)

  # Call create_heatmap function
  create_heatmap(
    gene = amino_acid_seq,
    order = unique(tsv_data$variants),
    chunk_size = 10,  # Adjust as needed
    scores_df = tsv_data,
    score_column = "am_pathogenicity",  # Adjust as needed
    variant_names = NULL,  # Use default variant names
    output_file = paste0(gene_name, "_heatmap.png")
  )
}


```

